Effects of the body force on the evacuation dynamics

### 2020-01-28 paper.tex
Agregue las correcciones de Dorso

We present in this section the results corresponding to the bottleneck geometry. We show the consequences of modifying the body force coefficient $k_n$ in the evacuation dynamics. Recall that this coefficient is associated with the stiffness of the pedestrians. \\

Fig.~\ref{vd_vs_te} shows the evacuation time as a function of the pedestrian's desired velocity. The evacuation time is defined as the time until 80\% of the pedestrians left the room. The different markers correspond to different $k_n$ values (\textit{i.e.} different body ``stiffness''; see caption for details). In this section we will focus on the evacuation time for $v_d>2\,$m/s.\\

\begin{figure}[htbp!]
\centering
\includegraphics[width=0.7\columnwidth]
{./vd_vs_te_N225.eps}
\caption{\label{vd_vs_te}Mean evacuation time (s) vs. the pedestrian’s desired velocity (m/s) for a bottleneck. The room was 20 m x 20 m size. The door was 0.92 m width. Mean values were computed from 10 evacuation processes. 225 pedestrians were initially placed in a square lattice with a random initial velocity. Each process was finished when 158 pedestrians left the room. The different symbols indicate the $k_n$ value corresponding to the body force (see the label). The crosses correspond to the Helbing original SFM parameter, the up-triangles correspond to the value measured in Ref.~\cite{gordon1989anthropometric}, squares correspond to zero body force and circles correspond to an extreme value of stiffness (one order of magnitude higher than the original SFM). The down triangles correspond to an intermediate value between the empirical value presented in Ref.~\cite{gordon1989anthropometric} and the chosen by Helbing in Ref.~\cite{helbing_2000} }
\end{figure}

Two behavioral patterns can be seen in Fig.~\ref{vd_vs_te}. Whenever the slope is positive, it means that the harder the pedestrians trying to get out (higher $v_d$), the higher it takes them to evacuate. This is the Faster-is-slower (FIS) regime. Conversely, the negative slope corresponds to a Faster-is-Faster (FIF) phenomenon (the harder they try, the quicker they leave). Fig.~\ref{vd_vs_te} shows a change in the slope sign only for stiffness values below $k=1.2$~E5. This means that ``soft'' individuals are required to attain this behavioral change.   Notice that for  $k_n=1.2$~E6 (the stiffest explored pedestrians), no FIS can be seen at all. Besides, the evacuation pattern for $k_n = 0$ and $k_n = 1.2$~E4 are very similar since the body force is comparable to the social force for these stiffness values. \\

@@

The regime for $v_d<2\,$m/s is different from the rest since most of the pedestrians are not touching each other. The only pedestrians who touch each other are the ones who are being re-injected in the room. The overlap is higher than, for example, $v_d=3\,$m/s because of the pedestrians who are re-injected bump against the bulk. \\

Besides, for a given $v_d$, the overlap increases as the $k_n$ value decreases. Reducing $k_n$ means reducing the normal body force, which allows a more significant invasion of the personal space of the individuals ( this means more overlap between pedestrians). The increment in mean overlap, however, seems to be more significant than the mean degree. \\

Increasing body stiffness has a significant impact on evacuation dynamics. The simulations reduction of the overlap and the mean degree between pedestrians, tend to diminish the sliding friction within the crowd. Consequently, an enhancement in the evacuation time (pedestrians escape faster) is observed, as pictured in Fig.~\ref{vd_vs_te}. Notice, however, that switching from a FIS regime (positive slope) to a  FIF regime (negative slope) in Fig.~\ref{vd_vs_te} appears as a more complex phenomenon. We will focus on this issue in an upcoming investigation.\\

Despite the sliding friction, the body force has a notorious impact in the number of pedestrians touching each other (say, the degree). This is clearly depicted in Fig.~\ref{network_bottleneck} where four different configurations of the evacuation dynamics are shown. The configurations represent 225 pedestrians trying to escape through a door see the caption for details. The colors correspond to the degree of each node (pedestrian), and the lines between pedestrians represent the contacts among them.\\

@@

In this research, we have explored the effects of the body force in the original SFM. We showed that the stiffness coefficient ($k_n$) has a significant impact on the evacuation dynamics (bottleneck) and also in the dynamics of pedestrians walking along a straight corridor.\\

In the bottleneck geometry, the evacuation time diminishes (pedestrians move faster) as pedestrians become stiffer for all the ranges of the desired velocities explored. This phenomenon happens because increasing the stiffness level promotes the repulsion between pedestrians pair interactions. As pedestrians detach from each other, the friction force diminishes due to the reduction of overlap and contacts among them. The general decrease in frictional force reduces the probability of producing a cluster of pedestrians blocking the exist (blocking cluster), which finally leads to a more efficient evacuation dynamics. \\

In the corridor geometry, the walking speed reduces as pedestrians become stiffer. This is the opposite behavior we obtained in bottleneck geometry. The difference is explained because although increasing stiffness further separates individuals, they can not detach from each other as much as possible due to the lateral walls (these walls are not relevant in the bottleneck geometry). Thus, when the stiffness increases, pedestrians reduce the overlap but increase the number of contacts (degree). In high density scenarios, the stiffness promotes the friction force.  Granular materials can behave like fluid or solids, depending on several conditions. If the stiffness level is low, the entire crowd behaves like a viscous fluid (with a parabolic velocity profile). However, if the stiffness level is high enough, the whole crowd behaves like a solid (with a constant velocity profile which depends on the friction against the walls).\\

>>>>

We present in this section the results corresponding to the bottleneck geometry. 
We show the consequences of modifying the body force coefficient $k_n$ in the 
evacuation dynamics. Recall that this coefficient is associated 
to the compression of the human body. \\

Fig.~\ref{vd_vs_te} shows the evacuation time as a function of the pedestrian's 
desired velocity for different values of $k_n$. The evacuation 
time is defined as the time until 80\% of the pedestrians have 
left the room. In this section we will focus on the evacuation time for 
$2\,$m/s$<v_d<10\,$m/s.\\

\begin{figure}[htbp!]
\centering
\includegraphics[width=0.7\columnwidth]
{./vd_vs_te_N225.eps}
\caption{\label{vd_vs_te}Mean evacuation time (s) vs. the pedestrian’s desired  
velocity (m/s) for a bottleneck. The room was 20 m x 20 m size. The door was 
0.92 m width. Mean values were computed from 10 evacuation processes. 225 
pedestrians were initially placed in a square lattice with a random initial 
velocity. Each process was finished when 158 pedestrians left the room. The 
different symbols indicate the $k_n$ value corresponding to the body force (see 
the label). The crosses correspond to the Helbing's original SFM parameter, the 
up-triangles correspond to the value measured in 
Ref.~\cite{anthropometric}, squares correspond 
to zero body force and circles correspond to an extreme value of stiffness (one 
order of magnitude higher than the original SFM). The 
 down triangles correspond to an intermediate value between the empirical value 
presented in Ref.~\cite{anthropometric} and the chosen by Helbing in 
Ref.~\cite{helbing_2000} }
\end{figure}

Two behavioral patterns can be seen in Fig.~\ref{vd_vs_te}. 
The region in which the slope is positive means that the 
harder the pedestrians try to get out (higher $v_d$), the 
longer it takes them to evacuate. This is the Faster-is-slower 
(FIS) regime. Conversely, the region in which the slope is 
negative  corresponds to a Faster-is-Faster (FIF) regime (the 
harder they try, the quicker they leave). Fig.~\ref{vd_vs_te} shows a change in 
in the behavior for desired velocities $v_d>1.2\,$m/s. The 
evacuation time is of type FIS+FIF for compression coefficients below 
$k=1.2\,$E5. This means that ``soft'' individuals are required to attain this 
behavioral change.  Notice that higher values of $k_n$ allow 
only FIS or FIF behaviors. For the highest explored value $k_n=1.2\,$E6, no 
FIS can be seen at all. Besides, the evacuation pattern for $k_n = 0$ and $k_n 
= 1.2\,$E4 are very similar since the body force is comparable to the social 
force for these stiffness values. \\

@@

The regime for $v_d<2\,$m/s is different from the rest since most of the 
pedestrians are not touching each other. The only pedestrians who touch each 
other are the ones who are being re-injected in the room. The overlap is higher 
than, for example, $v_d=3\,$m/s because of the pedestrians that 
are re-injected and collide with pedestrians in the bulk. \\

Besides, for a given $v_d$, the overlap increases as the $k_n$ value decreases. 
Reducing $k_n$ means reducing the normal body force, which allows a more 
significant intrusion of the personal space of the individuals 
(this means more overlap between pedestrians).\\

Increasing the body stiffness has a significant impact on 
evacuation dynamics. The reduction of the overlap and the mean degree between 
pedestrians in the simulation, tend to diminish the sliding 
friction within the crowd. Consequently, an enhancement in the evacuation time 
(pedestrians escape faster) is observed, as pictured in Fig.~\ref{vd_vs_te}. 
Notice, however, that switching from a FIS regime (positive slope) to a  FIF 
regime (negative slope) in Fig.~\ref{vd_vs_te} appears as a more complex 
phenomenon. We will focus on this issue in an upcoming investigation.\\

Despite the effects on the sliding friction, the body force 
has a notorious impact in the number of pedestrians touching each other (say, 
the degree). This is clearly depicted in Fig.~\ref{network_bottleneck} where 
four different configurations of the evacuation dynamics are shown. The 
configurations represent 225 pedestrians trying to escape through a door see the 
caption for details. The colors correspond to the degree of each node 
(pedestrian), and the lines between pedestrians represent the contacts among 
them.\\

@@

In this research, we explore the effect on the pedestrians dynamics of the (sometimes neglected) body force in the framework of the SFM. We showed that the stiffness coefficient ($k_n$) has a significant impact on the evacuation dynamics (bottleneck) and also in the dynamics of pedestrians walking along a straight corridor.\\

In the bottleneck geometry, the evacuation time diminishes (pedestrians move faster) as pedestrians become stiffer for all the ranges of the desired velocities explored. This phenomenon occurs because increasing the stiffness level promotes the repulsion between pedestrians pair interactions. As pedestrians detach from each other, the friction force diminishes due to the reduction of overlap between pedestrians and the contacts among them. The general decrease in frictional force reduces the probability of producing a cluster of pedestrians blocking the exist (blocking cluster), which leads to a more efficient evacuation dynamics. \\

In the corridor geometry, the actual walking speed reduces as pedestrians become stiffer. This is the opposite behavior with respect to the obtained in bottleneck geometry. The difference is explained because although increasing stiffness further separates individuals, they can not detach from each other as much as possible due to the lateral walls (these walls are not relevant in the bottleneck geometry). Thus, when the stiffness increases, the overlap between pedestrians is reduced while the number of contacts (degree) increases. Granular materials can behave like fluid or solids, depending on several conditions. If the stiffness level is low, the entire crowd behaves like a viscous fluid (with a parabolic velocity profile). However, if the stiffness level is high enough, the whole crowd behaves like a crystalline solid with a constant velocity profile which only depends on the friction interaction with the walls.\\
